Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2024-10-04 Latex_files/Thesis/chapters/chapter 1/main.tex
updates 04-10-2024

\chapter{Introduction}
Crystallisation is one of the oldest forms of phase separation used by humanity\cite{Schoen1956}, put simply, it is the formation and growth 
of a new structured phase within a disordered bulk phase. This has 
applications in a number of industries such as pharmaceuticals, food 
production, and electronics \cite{Myerson2002}. Where the extraction 
of dilute materials can help improve product quality while keeping 
production costs low. The industrialisation of crystallisation has 
allowed engineers to reliably and efficiently induce the crystal 
formation within a bulk phase. However, while on a large scale 
crystallisation is seemingly an understood physical process, just a 
small amount of investigation into the literature reveals that at a 
micro scale there is yet to be unifying theory that can accurately 
explain the process of crystallisation \cite{Fu2021}. Of which there 
two primary areas of focus, nucleation and crystal growth. The latter 
focusing on how an already stable crystal grows and how it takes on its 
final shape. Whereas the former is more concerned with which factors 
contribute to its formation in order to control crystal growth.

The focus of this research is partly based around investigating the impact 
of a focused electromagnetic (EM) field can influence the nucleation process 
by inducing shear flow within the bulk phase. This in hope that the 
localised EM field can be used to study different phenomena in nucleation 
in order to develop a better understanding of the underlying mechanisms.

@@

\subsection{Primary \& Secondary nucleation}
From an industrial perspective, the nucleation process can be broadly 
categorised into either primary of secondary nucleation. The former 
describes the formation of an initial nucleus within the bulk phase, 
absent of any external stimuli. Primary nucleation is therefore considered 
stochastic as there is no reliable means of predicting where a nucleus
may form, or how long it will take. The only reliable indicator being 
that higher supersaturations will result in faster nucleation rates. At 
a small scale one can estimate the nucleation rate by making repeated 
measurements of sample solutions and seeing how many have nucleated 
after a given time, giving us a Poisson probability distribution.

\begin{align}
	P(t) = 1 - exp\left[-JV(t-t_g)\right] = \frac{M^*(t)}{M}
\end{align}

@@

In contrast, secondary nucleation is the result of a initial 
seed crystal inducing further nucleation within the bulk solution
\cite{Botsaris1976}, in absence of a seed crystal nucleation is
near impossible. The research of Secondary Nucleation is fascinating
for its own merits; is it speculated that the supersaturation 
barrier is non-existent \cite{Cashmore2022}, and furthermore there 
are a multitude of mechanisms that can induce secondary 
nucleation making it imperative that researchers can control the
nucleation rate at an industrial scale.

\begin{figure}[h!]
	\includegraphics[width=\linewidth]{secondary_nucleation.jpg}
	\caption{Secondary Nucleation mechanisms, classified by Agrawal and Paterson \cite{Agrawal2015}}
\end{figure}

\section{Nucleation Theories}
\subsection{Classical Nucleation Theory (CNT)}
Sometimes referred to as 'Gibs Nucleation Theory' the original theory 
was first formed from the works of Volmer and Weber, and Frenkel 
\cite{Frenkel1939, Volmer1926}. While initially it was more focused 
into describing droplet formation in condensing vapours it was 
extrapolated to describe crystallisation. The central premise of 
classical theory is that nucleation occurs stochastically due to 
collisions between individual solute molecules, ions, or atoms. At the 
same time the bulk phase is resistant to the formation of a new phase. 
The competition between these random collisions and the bulk solution 
can be used to predict the probability of a newly formed nucleus.

@@

\subsection{Non-classical Nucleation}

The current research into two step nucleation (or now more commonly referred 
to as multi-step nucleation theory) is developing a robust framework to 
describe what nucleation pathway will occur given the initial conditions. 
Reviews of all currently documented nucleation pathways identified highlighted 
the fact the need for a the development of \textit{in situ} techniques that 
can induce nucleation locally but can also reliably identify the nucleation 
pathway across a broad range of experimental conditions \cite{Karthika2016, Fu2021}.

@@

\subsection{Gaussian Beams}
A Gaussian Beam is a simplistic approximation of a laser beam, in short
it is an unbounded plane wave whose intensity of the electromagnetic 
field falls off from the centre similar to a normal distribution. This 
is perfectly fine for beams with a divergence angle of $0^{\circ}$, but 
later studies into focused beams has shown that the assumption of a 
Gaussian intensity profile is inaccurate. A study conducted by Lock and Gouesbet found that as the Gaussian beam was focused the fall off in 
the electric field would consequentially induce additional electromagnetic fields \cite{Lock1994}. As a result the shape of the beam profile diverges from a Gaussian profile for focused beams. Regardless, for optical trapping experiments it is useful to approximate a Gaussian beam as many descriptions
of lasers are simply combinations of plane waves (with varying amplitudes
and wave vectors). More information on electromagnetic theory is covered in 
chapter 2 but for now any mention of a focused laser can be thought of as a
Gaussian beam, unless specified otherwise.

@@

\section{Significance of Thesis}
As I have hoped to make clear in the above introduction, the current 
state nucleation theory is rather cumbersome at a micro-level. Models
such as CNT and multi-step nucleation are not sufficient for describing
the myriad of potential pathways nucleation can go down. As suggested
by some review articles, the best way to address this is by developing 
\textit{in-situ} methods that can study the pre nucleation phase in 
greater detail \cite{Fu2021}. Furthermore, the ability to localise 
nucleation allows for better characterisation of the kinetics of crystal 
growth. Laser induced nucleation stands to be an ideal method to study
nucleation, as the laser output can be concentrated to a small area. 
In this project we want to consider another application of optical 
tweezers, their ability to rotate micro-rotors to generate fluid flow.
Such a method has not been tried before and would allow for us to 
consider how fluid flow impacts nucleation dynamics at the micro-level.

@@

In addition, computer simulations studying the dynamics of asymmetric dimers 
and their applications as complex micro-rotors are discussed in chapter 4. 
The latter results are interesting as a potential insight into how optical
torque is transferred to anisotropic objects.

Lastly two novel methods of measuring rotational motion are discussed in 
chapter 5; firstly via a novel detection fibre method that allows for 
instantaneous measurements of the orientational behaviour of optically 
trapped ellipsoids/dimers; and secondly we create a simulative quadrant 
photo diode that replicates laboratory results, utilising linear regression 
techniques we measure the change in orientation in order to measure the 
optical torque applied to a non-birefringent particle.

>>>>

\chapter{Introduction}
Crystallisation is one of the oldest forms of phase separation used by humanity\cite{Schoen1956}, put simply, it is the formation and growth 
of a new structured phase within a disordered bulk phase. This has 
applications in a number of industries such as pharmaceuticals, food 
production, and electronics \cite{Myerson2002}. Where the extraction 
of dilute materials can help improve product quality while keeping 
production costs low. The industrialisation of crystallisation has 
allowed engineers to reliably and efficiently induce the crystal 
formation within a bulk phase. However, while on a large scale 
crystallisation is seemingly an understood physical process, just a 
small amount of investigation into the literature reveals that at a 
micro scale there is yet to be unifying theory that can accurately 
explain the process of crystallisation \cite{Fu2021}. Of which there 
two primary areas of focus, nucleation and crystal growth. The latter 
focusing on how an already stable crystal grows and how it takes on its 
final shape. Whereas the former is more concerned with which factors 
contribute to its formation in order to control crystal growth.SW

In this chapter we will outline our current understanding of nucleation,
why there is a gap in the literature, and why there is still a need for
local control of nucleation events. Furthermore, we will also highlight
recent developments involving the use of optical tweezers (and other 
laser based methods) in order to develop our understanding of nucleation.
Lastly we highlight the connection between optical tweezer's ability to 
probe fluid viscosity via rotational motion, and the known link between
nucleation and fluid flow, to propose a novel potential method of 
creating localised nucleation events.

@@

\subsection{Primary \& Secondary nucleation}
From an industrial perspective, the nucleation process can be broadly 
categorised into either primary of secondary nucleation. The former 
describes the formation of an initial nucleus within the bulk phase, 
absent of any external stimuli. Primary nucleation is therefore considered 
stochastic as there is no reliable means of predicting where a nucleus
may form, or how long it will take. The only reliable indicator being 
that higher supersaturations will result in faster nucleation rates. At 
a small scale one can estimate the nucleation rate by making repeated 
measurements of sample solutions and seeing how many have nucleated 
after a given time, giving us a Poisson probability distribution.
\begin{align}
	P(t) = 1 - exp\left[-JV(t-t_g)\right] = \frac{M^*(t)}{M}
\end{align}

@@

In contrast, secondary nucleation is the result of a initial 
seed crystal inducing further nucleation within the bulk solution
\cite{Botsaris1976}. The research of Secondary nucleation is 
fascinating for its own merits; is it speculated that the 
supersaturation barrier is non-existent \cite{Cashmore2022}, 
Furthermore control of secondary nucleation events are crucial 
for ensuring reliable industrial crystalliser performance, 
fig.~\ref{fig:secondary} depicts an attempt of classifying  
every possible mechanism that could lead to secondary nucleation.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.95\linewidth]{secondary_nucleation.jpg}
	\caption{Secondary Nucleation mechanisms, classified by Agrawal 
		and Paterson \cite{Agrawal2015}}
	\label{fig:secondary}
\end{figure}

Its clear that preventing secondary nucleation is impossible
with so many possible mechanism, modelling its contribution
to the nucleation rate has been the area of research for some
time now \cite{Cashmore2022, Flannigan2023}. In most cases 
secondary nucleation is actually desirable for large scale 
production as primary nucleation is a stochastic process.

\section{Nucleation Theories}
A number of potential theories have been proposed to explain 
how nucleation occurs at a microscopic scale. In doing so 
we could potentially predict the expected yield of a given
crystalliser based on the initial conditions of the 
solution. Outlined below are some of the most popular theories
currently used to describe nucleation.

\subsection{Classical Nucleation Theory (CNT)}
Sometimes referred to as 'Gibs Nucleation Theory' the original theory 
was first formed from the works of Volmer and Weber, and Frenkel 
\cite{Frenkel1939, Volmer1926}. While initially it was more focused 
into describing droplet formation in condensing vapours it was 
extrapolated to describe crystallisation. The central premise of 
classical theory is that nucleation occurs stochastically due to 
collisions between individual solute molecules, ions, or atoms. At the 
same time the bulk phase is resistant to the formation of a new phase. 
The competition between these random collisions and the bulk solution 
can be used to predict the probability of a newly formed nucleus.

@@

\subsection{Non-classical Nucleation} 
The current research into two step nucleation (or now more commonly 
referred to as multi-step nucleation theory) is developing a robust 
framework to describe what nucleation pathway will occur given the 
initial conditions. Reviews of all currently documented nucleation 
pathways identified highlighted the fact the need for a the 
development of \textit{in situ} techniques that can induce nucleation 
locally but can also reliably identify the nucleation pathway across 
a broad range of experimental conditions \cite{Karthika2016, Fu2021}.

@@

\subsection{Gaussian Beams}
A Gaussian Beam is a simplistic approximation of a laser beam, in 
short it is an unbounded plane wave whose intensity of the 
electromagnetic field falls off from the centre similar to a normal distribution. This is perfectly fine for beams with a divergence 
angle of $0^{\circ}$, but later studies into focused beams showed 
that the assumption of a Gaussian intensity profile is inaccurate. 
A study conducted by Lock and Gouesbet found that as the Gaussian 
beam was focused the fall off in the electric field would 
consequentially induce additional electromagnetic fields 
\cite{Lock1994}. As a result the shape of the beam profile diverges 
from a Gaussian profile for focused beams. Regardless, for optical 
trapping experiments it is useful to assume a near Gaussian beam as 
many descriptions of lasers are simply combinations of plane waves 
(with varying amplitudes and wave vectors).  More information on 
electromagnetic theory is covered in chapter 2 but for now any 
mention of a focused laser can be thought of as a Gaussian beam, 
unless specified otherwise.

@@

\section{Significance of Thesis}
As I have hoped to make clear in the above introduction, the 
current state nucleation theory is rather cumbersome at a 
micro-level. Models such as CNT and multi-step nucleation are 
not sufficient for describing the myriad of potential pathways 
nucleation can go down. As suggested by some review articles, 
the best way to address this is by developing \textit{in-situ} 
methods that can study the pre nucleation phase in greater detail \cite{Fu2021}. Furthermore, the ability to localise nucleation 
allows for better characterisation of the kinetics of crystal 
growth. Laser induced nucleation stands to be an ideal method to 
study nucleation, as the laser output can be concentrated to a 
small area \cite{Korede2023}. This fine control over the local 
fluid would allow for individual nucleation events to be
characterised as a factor of the local fluid properties. One 
challenge lies in the fact that its clear that the local fluid 
properties have a direct influence on the likelihood of laser
induced nucleation from occurring \cite{Korede2023, Ward2016,
Yuyama2012, Liao2022}.

A way around this would be to try and induce the local fluid 
without directly relying on the electromagnetic field. Optical 
tweezers can reliably do so already by applying an optical 
torque to a trapped particle in order to induce fluid flow 
\cite{Bishop2004, RobertsonAnderson2018}. It's already well
documented that shearing will enhance nucleation events at a
macro-level \cite{Debuysschere2023}. Micro-rheological studies 
using optical tweezing have been more interested in probing 
the local viscosity rather than try and use it as a means of 
shearing the fluid to induce crystal growth.

@@

The latter chapters cover computer simulations into the 
behaviour of microscopic spherical dimers in an optical
trap. Prior research into dimers using back focal plane
interferometry has mostly considered their trapping 
behaviour to be similar to a single sphere but with a 
difference in trapping strength. Computer simulations
reveal a host of new behaviours dependent on the dimer's
size, orientation, proximity to the trapping focus, and 
even the polarisation of the trapping beam. The latter in
particular suggests that multi-spherical particles can act
as sophisticated micro-rotors.

However, this raises a its own host of experimental challenges,
namely how do we characterise the behaviour of an arbitrary
particle. Relying on current characterisation techniques is not
possible as they are predicated on the trapping object to behave
like an isolated sphere. Two novel methods of measuring rotational 
motion are discussed in chapter 5; firstly via a novel detection 
fibre method that allows for instantaneous measurements of the 
orientational behaviour of optically trapped ellipsoids/dimers; 
and secondly we create a simulative quadrant photo diode that 
replicates laboratory results, utilising linear regression 
techniques we measure the change in orientation in order to 
measure the optical torque applied to a non-birefringent particle.
